---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

Loading Libraries and changing default
```{r}
 # This accounts for a limitation in RStudio
  options("max.print" = 9999999)
  
  #load libraries needed
  library(ggplot2)
  library(grid)
  library(dplyr)
 
```

## Loading and preprocessing the data

The data were downloaded from [The Course Website](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip)

These data were then loaded into the current working directory (as a csv file), activity.csv
A dataframe, activity, was generated via read.csv

```{r}
activity <- read.csv("~\\R\\datasciencecoursera\\RepData_PeerAssessment1\\activity.csv")
str(activity)
summary(activity)
```


## What is mean total number of steps taken per day?

```{r stepsperday}
 
  #load data into function
  activity <- read.csv("~\\R\\datasciencecoursera\\RepData_PeerAssessment1\\activity.csv")
  #remove NA
  stepdata <- activity[!is.na(activity$steps),]
  #sum step data by day
     newstep <- aggregate(steps ~ date, data = stepdata, sum)
     newstepmean <- sprintf("%.4f",mean(newstep$steps))
     newstepmedian <- median(newstep$steps)
     
     #plot
     ggplot(data = newstep, aes(x = steps)) + geom_histogram() +
       labs(title = "Distribution of Total Steps per Day from Watch Data") +
       xlab("Total Steps per Day") +
       ylab("Frequency of step count")
       
     #need to fix x-axis
```

Participants walk an average of `r newstepmean` daily.

The median number of steps participants take daily is `r newstepmedian`.


## What is the average daily activity pattern?
A time-series plot (type = "l") of the five minute interval (x-axis) and the average number of steps taken averaged across all days (y-axis).
Demonstrate which interval, on average across all days in the dataset, contains the maximum number of steps.

```{r steppattern}

  #load data into function
  activity <- read.csv("~\\R\\datasciencecoursera\\RepData_PeerAssessment1\\activity.csv")
  #remove NA
  stepdata <- activity[!is.na(activity$steps),]
  #sum step data by day
  newstep <- aggregate(steps ~ interval, data = stepdata, mean)
  maxsteps <- newstep$interval[newstep$steps == max(newstep$steps)]
  
  #plot
  ggplot(data = newstep) + geom_point(mapping= aes(x = interval, y = steps)) +
    labs(title = "Average Steps per Interval from Watch Data") +
    xlab("Time Interval") +
    ylab("Average Steps per Interval")
  
```
  
  The time interval where the participant mean step count is the highest is `r maxsteps`.


## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
